Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2020-10-09 minimax.tex
minors spellCheck

Let $V$ be the set of possible preference profiles (the union, for any integer $n$, of the $n$-fold cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the emptyset (sets are used for tied winners).
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
	\label{eq:srule}
	s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
	= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with the highest score.

@@

Let $(x^{*},\bar{y}, \bar{\profile}, \bar{\w})$ be the current solution of the minimax regret, where $\bar{y}, \bar{\profile}, \bar{\w}$ are, respectively, the choice of $y$, the completion of the profile and the selection of weights for which $x^{*}$ is the minimax optimal alternative. 
The \strat{Pessimistic} strategy considers a set of $n + m$ candidate questions: one per agent, and one per rank.
The candidate questions to the agents use the heuristic proposed by \citet{Lu2011}: for each agent $j$, if $x^*$ and $\bar{y}$ are incomparable in $\ppref_j$, the candidate question concerns the pair $(x^*, \bar{y})$, otherwise, the candidate question concerns the pair $(x^*, z)$ for some $z$ incomparable to $x^*$ (randomly chosen), or if none such $z$ exist, the pair $(\bar{y}, z)$ for some $z$ incomparable to $\bar{y}$, or, if both $x^*$ and $\bar{y}$ are comparable to every alternatives in $\ppref_j$, any incomparable pair is picked at random. 
To understand the intuition behind favoring these two alternatives, recall that, as shown in the proof of \cref{claim:completion}, in order to increase $\PMR(x^{*},\bar{y})$, an adversary should place as many alternatives as she can above $x^{*}$ if $x^* \ppref_j \bar{y}$, and between $\bar{y}$ and $x^{*}$ if $\bar{y} \ppref_j x^*$.  Asking questions involving $x^*$ or $\bar{y}$ thus has a particular chance of reducing usefully the possibilities of the adversary.
The candidate questions to the chair are determined as in the Random strategy.
The strategy then selects among these $n + m$ candidate questions one that leads to minimal regret in the worst case, considering both possible answers to the question, and with penalty terms depending on the kind of question. To define this comparison precisely, assume that a question $q_1$ has type $t_1$ (being “chair” or “agent”), and leads to the possible new knowledge states $(\pprofile_1, W_1)$ and $(\pprofile'_1, W'_1)$, depending on the answer. 
Define $R^{\max}_1 = \max\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)}$
and $R^{\min}_1 = \min\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)} p(t) + p'(t)$.
%If $(\pprofile_1, W_1) ≥ (\pprofile'_1, W'_1)$, define $\MMR^{\max} = \MMR(\pprofile_1, W_1)$ and $\MMR^{\min} = \MMR(\pprofile'_1, W'_1)$; otherwise, define $\MMR^{\max} = \MMR(\pprofile'_1, W'_1)$ and $\MMR^{\min} = \MMR(\pprofile_1, W_1)$.
% and and $(\pprofile^\min_2, W^\min_2)$. numbering them so that $\MMR(\pprofile_1, W_1) ≥ \MMR(\pprofile_2, W_2)$. Then the badness of the question in the worst case is:
The terms $p(t)$ and $p'(t)$ are real numbers depending on the kind of question. These parameters are to be determined experimentally.
Define similarly $t_2$, $R^{\max}_2$ and $R^{\min}_2$ for a question $q_2$.
Then, the Pessimistic strategy considers question $q_1$ more suitable than $q_2$ iff
\begin{align}
	R^{\max}_1 ≤ R^{\max}_2 \text{ or } [R^{\max}_1 = R^{\max}_2 \text{ and } R^{\min}_1 ≤ R^{\min}_2].
\end{align}
%This comparison gives a way of picking questions among a set of possible questions, by picking one that minimizes this approximate measure of minimax regret {\em a posteriori}. 
We use this operator, similar to a $\max$ operator, because \citet{Cailloux2014} reported that pessimistic aggregation works much better than optimistic aggregation, although our experiments in this context suggested a weak impact of that choice on the performance of the strategy.
However, to avoid the absorption property of max, we adopt a criterion closer to $\leximax$ as an aggregator: if the maximal MMR of two questions are (nearly) equal, then it considers the penalized minimal MMR values, preferring the question with the lowest value.
(Technically, instead of applying a pure lexicographic aggregation, which is very sensitive to small errors due to floating-point computations, we apply a weighted sum between the maximal and the penalized minimal MMR values, with a predominant weight to the maximal one.)

The \strat{Extended pessimistic} strategy uses the same criterion as the pessimistic strategy, but extending it to a bigger set of candidate questions, the same as those considered by the Random strategy.
These candidate questions are then evaluated using the same operator as for the Pessimistic strategy, that is, we compute the minimax regret for the two possible answers, for each candidate question, and choose the best according to (quasi-)$\leximax$.
We use this strategy to test whether Pessimistic performs well: depending on the quality of the heuristic of the Pessimistic strategy, it might perform nearly equally well, or perhaps even better, than Extended pessimistic, while being (much) faster. Extended pessimistic is applicable only to small problem instances: its complexity is in $O(n^2 m^5)$, because we consider $O(m^2)$ questions for each agent and need for each question to compute MMR twice, whose complexity is $O(nm^3)$.

@@

We start with empty knowledge ($\pprofile = (\emptyset, \emptyset, …), W = \W$) about the preference orderings of the agents or the weight differences favored by the chair. We obtain the first question to be asked using the strategy under test, as described above. We use the oracle to answer the question and update our knowledge, which is thus used to obtain the next question. This is repeated until $k$ answers have been obtained, compute the resulting MMR values along the way for various values of $k$. We repeat this whole experiment a variable number of times indicated below, for a given $(m, n, k)$, and report the average resulting MMR and the standard deviation \textit{sd}.

@@

We mention some directions for future works.
Further development of elicitation strategies, considering alternative heuristics, is an important direction. 
Second, elicitation could be extended to voting rules beyond scoring rules. 
Finally, an important direction of extension aims at studying how to elicit preferences while restraining to concrete and easy questions.
% Acknowledgements: We thank the reviewers for comments helping to improve the paper.

>>>>

Let $V$ be the set of possible preference profiles (the union, for any integer $n$, of the $n$-fold Cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the empty set (sets are used for tied winners).
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$ (also called the scoring vector).
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
	\label{eq:srule}
	s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
	= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with the highest score.

@@

Let $(x^{*},\bar{y}, \bar{\profile}, \bar{\w})$ be the current solution of the minimax regret, where $\bar{y}, \bar{\profile}, \bar{\w}$ are, respectively, the choice of $y$, the completion of the profile and the selection of weights for which $x^{*}$ is the minimax optimal alternative. 
The \strat{Pessimistic} strategy considers a set of $n + m$ candidate questions: one per agent, and one per rank.
The candidate questions to the agents use the heuristic proposed by \citet{Lu2011}: for each agent $j$, if $x^*$ and $\bar{y}$ are incomparable in $\ppref_j$, the candidate question concerns the pair $(x^*, \bar{y})$, otherwise, the candidate question concerns the pair $(x^*, z)$ for some $z$ incomparable to $x^*$ (randomly chosen), or if none such $z$ exist, the pair $(\bar{y}, z)$ for some $z$ incomparable to $\bar{y}$, or, if both $x^*$ and $\bar{y}$ are comparable to every alternatives in $\ppref_j$, any incomparable pair is picked at random. 
To understand the intuition behind favoring these two alternatives, recall that, as shown in the proof of \cref{claim:completion}, in order to increase $\PMR(x^{*},\bar{y})$, an adversary should place as many alternatives as she can above $x^{*}$ if $x^* \ppref_j \bar{y}$, and between $\bar{y}$ and $x^{*}$ if $\bar{y} \ppref_j x^*$.  Asking questions involving $x^*$ or $\bar{y}$ thus has a particular chance of reducing usefully the possibilities of the adversary.
The candidate questions to the chair are determined as in the Random strategy.
The strategy then selects among these $n + m$ candidate questions one that leads to minimal regret in the worst case, considering both possible answers to the question, and with penalty terms depending on the kind of question. To define this comparison precisely, assume that a question $q_1$ has type $t_1$ (being “chair” or “agent”), and leads to the possible new knowledge states $(\pprofile_1, W_1)$ and $(\pprofile'_1, W'_1)$, depending on the answer. 
Define $R^{\max}_1 = \max\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)}$
and $R^{\min}_1 = \min\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)} p(t) + p'(t)$.
%If $(\pprofile_1, W_1) ≥ (\pprofile'_1, W'_1)$, define $\MMR^{\max} = \MMR(\pprofile_1, W_1)$ and $\MMR^{\min} = \MMR(\pprofile'_1, W'_1)$; otherwise, define $\MMR^{\max} = \MMR(\pprofile'_1, W'_1)$ and $\MMR^{\min} = \MMR(\pprofile_1, W_1)$.
% and and $(\pprofile^\min_2, W^\min_2)$. numbering them so that $\MMR(\pprofile_1, W_1) ≥ \MMR(\pprofile_2, W_2)$. Then the badness of the question in the worst case is:
The terms $p(t)$ and $p'(t)$ are real numbers depending on the kind of question. These parameters are to be determined experimentally.
Define similarly $t_2$, $R^{\max}_2$ and $R^{\min}_2$ for a question $q_2$.
Then, the Pessimistic strategy considers question $q_1$ more suitable than $q_2$ iff
\begin{align}
	R^{\max}_1 ≤ R^{\max}_2 \text{ or } [R^{\max}_1 = R^{\max}_2 \text{ and } R^{\min}_1 ≤ R^{\min}_2].
\end{align}
%This comparison gives a way of picking questions among a set of possible questions, by picking one that minimizes this approximate measure of minimax regret {\em a posteriori}. 
We use this operator, similar to a $\max$ operator, because \citet{Cailloux2014} reported that pessimistic aggregation works much better than optimistic aggregation, although our experiments in this context suggested a weak impact of that choice on the performance of the strategy.
However, to avoid the absorption property of max, we adopt a criterion closer to $\leximax$ as an aggregator: if the maximal $\MMR$ of two questions are (nearly) equal, then it considers the penalized minimal $\MMR$ values, preferring the question with the lowest value.
(Technically, instead of applying a pure lexicographic aggregation, which is very sensitive to small errors due to floating-point computations, we apply a weighted sum between the maximal and the penalized minimal $\MMR$ values, with a predominant weight to the maximal one.)

The \strat{Extended pessimistic} strategy uses the same criterion as the pessimistic strategy, but extending it to a bigger set of candidate questions, the same as those considered by the Random strategy.
These candidate questions are then evaluated using the same operator as for the Pessimistic strategy, that is, we compute the minimax regret for the two possible answers, for each candidate question, and choose the best according to (quasi-)$\leximax$.
We use this strategy to test whether Pessimistic performs well: depending on the quality of the heuristic of the Pessimistic strategy, it might perform nearly equally well, or perhaps even better, than Extended pessimistic, while being (much) faster. Extended pessimistic is applicable only to small problem instances: its complexity is in $O(n^2 m^5)$, because we consider $O(m^2)$ questions for each agent and need for each question to compute $\MMR$ twice, whose complexity is $O(nm^3)$.

@@

We start with empty knowledge ($\pprofile = (\emptyset, \emptyset, …), W = \W$) about the preference orderings of the agents or the weight differences favored by the chair. We obtain the first question to be asked using the strategy under test, as described above. We use the oracle to answer the question and update our knowledge, which is thus used to obtain the next question. This is repeated until $k$ answers have been obtained, compute the resulting $\MMR$ values along the way for various values of $k$. We repeat this whole experiment a variable number of times indicated below, for a given $(m, n, k)$, and report the average resulting $\MMR$ and the standard deviation \textit{sd}.

@@

We mention some directions for future works.
Further development of elicitation strategies, considering alternative heuristics, is an important direction. 
Second, elicitation could be extended to voting rules beyond scoring rules. 
Finally, an important direction of extension aims at studying how to elicit preferences while restraining to concrete and easy questions.
% Acknowledgments: We thank the reviewers for comments helping to improve the paper.
